\Large\textbf{Advection versus diffusion in brain ventricular transport}

### 2025-04-11 paper/main.tex
abstract

\begin{abstract}
Cerebrospinal fluid (CSF) is integral to brain function, as the fluid
provides mechanical support for the brain and the flow of CSF helps
distribute nutrients, neurotransmitters and metabolites throughout the
central nervous system. CSF flow is partly driven by beating of motile
cilia, hair-like structures that line the ependymal cell layer of the
walls of the brain ventricles. Despite the physiological importance of
CSF flow, the underlying mechanisms of cilia-mediated flow and
transport in the brain ventricles remain yet to be comprehensively
resolved. This study analyzes and evaluates the role of motile cilia
in brain ventricular CSF flow and transport. The geometry of the
zebrafish larval brain ventricles, for which detailed knowledge of
cilia properties and CSF motion velocities and frequencies exists, was
used to develop finite element methods for modelling flow and
transport. The computational model is validated by in vivo experiments
that monitor transport of a photoconvertible protein secreted in the
brain ventricles. Our results show that while cilia contribute to
advection of large particles, diffusion plays a significant role in
the transport of small solutes. We also demonstrate how cilia location
and the geometry of the ventricular system impact solute
distribution. Altogether, this work presents a computational framework
that can be applied to other ventricular systems, together with new
concepts of how molecules are transported within the brain and its
ventricles.

>>>>

\begin{abstract}
Cerebrospinal fluid (CSF) is integral to brain function. CSF
provides mechanical support for the brain and helps
distribute nutrients, neurotransmitters and metabolites throughout the
central nervous system. CSF flow is driven by several processes,
including the beating of motile cilia located on the
walls of the brain ventricles. Despite the physiological importance of
CSF, the underlying mechanisms of CSF flow and solute transport
in the brain ventricles remain to be comprehensively
resolved. This study analyzes and evaluates specifically
the role of motile cilia in CSF flow and transport. 
We developed finite element methods for modeling flow
and transport using the geometry of the zebrafish larval
brain ventricles, for which we have detailed knowledge of
cilia properties and CSF motion.
The computational model is validated by in vivo experiments
that monitor transport of a photoconvertible protein secreted in the
brain ventricles. Our results show that while cilia contribute to
advection of large particles, diffusion plays a significant role in
the transport of small solutes. We also demonstrate how cilia location
and the geometry of the ventricular system impact solute
distribution. Altogether, this work presents a computational framework
that can be applied to other ventricular systems, together with new
concepts of how molecules are transported within the brain and its
ventricles.
